Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-25 paper.qmd
Updated trimesh for pbmc tSNE model

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = umap_pbmc_scaled,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  q = 0.05,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC_"
)

@@

trimesh_removed_pbmc <- vis_rmlg_mesh(
  distance_edges = distance_pbmc,
  benchmark_value = benchmark_pbmc,
  tr_coord_df = tr_from_to_df_pbmc,
  distance_col = "distance") +
  geom_point(
    data = umap_pbmc_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.3,
    size = 0.1,
    color = "#000000"
  )

## Compute error
error_df <- augment(
  df_bin_centroids = df_bin_centroids_pbmc,
  df_bin = df_bin_pbmc,
  training_data = training_data_pbmc,
  newdata = NULL,
  type_NLDR = "UMAP",
  col_start = "PC_")

error_df <- error_df |>
  mutate(type = case_when(
    row_wise_abs_error <= 5 ~ "error 0-5",
    row_wise_abs_error <= 10 ~ "error 5-10",
    row_wise_abs_error <= 15 ~ "error 10-15",
    row_wise_abs_error <= 20 ~ "error 15-20",
    .default = "error greter than 20"
  )) |>
  mutate(type = factor(type, levels = c(
    "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error greter than 20")))

## To join embedding
error_df <- error_df |>
  bind_cols(umap_pbmc_scaled |>
              select(-ID))

error_plot_pbmc <- error_df |>
  ggplot(aes(x = UMAP1,
             y = UMAP2,
             color = type,
             group = ID)) +
  geom_point(alpha=0.5,
             size = 0.1) +
  interior_annotation("b", sc_ltr_pos)

@@

>>>>

pbmc_model <- fit_highd_model(
  training_data = training_data_pbmc,
  emb_df = umap_pbmc_scaled,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  q = 0.1,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC_"
)

@@

tr_df <- distinct(tibble(
  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))

distance_df_small_edges_pbmc <- distance_pbmc |>
  filter(distance < benchmark_pbmc)

tr_from_to_df_pbmc <- inner_join(
  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
  by = c("from", "to"))

# trimesh_removed_pbmc <- vis_rmlg_mesh(
#   distance_edges = distance_pbmc,
#   benchmark_value = benchmark_pbmc,
#   tr_coord_df = tr_from_to_df_pbmc,
#   distance_col = "distance") +
#   geom_point(
#     data = umap_pbmc_scaled,
#     aes(
#       x = UMAP1,
#       y = UMAP2
#     ),
#     alpha = 0.3,
#     size = 0.1,
#     color = "#000000"
#   )

trimesh_removed_pbmc <- ggplot() + 
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#33a02c") +
  geom_point(
    data = umap_pbmc_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.3,
    size = 0.1,
    color = "#000000"
  )  +
  interior_annotation("a")

# ## Compute error
# error_df <- augment(
#   df_bin_centroids = df_bin_centroids_pbmc,
#   df_bin = df_bin_pbmc,
#   training_data = training_data_pbmc,
#   newdata = NULL,
#   type_NLDR = "UMAP",
#   col_start = "PC_")

# ## Categorize error
# 
# error_df <- error_df |>
#   mutate(type = case_when(
#     row_wise_abs_error <= 5 ~ "error 0-5",
#     row_wise_abs_error <= 10 ~ "error 5-10",
#     row_wise_abs_error <= 15 ~ "error 10-15",
#     row_wise_abs_error <= 20 ~ "error 15-20",
#     .default = "error greter than 20"
#   )) |>
#   mutate(type = factor(type, levels = c(
#     "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error greter than 20")))
# 
# ## To join embedding
# error_df <- error_df |>
#   bind_cols(umap_pbmc_scaled |>
#               select(-ID))
# 
# error_plot_pbmc <- error_df |>
#   ggplot(aes(x = UMAP1,
#              y = UMAP2,
#              color = type,
#              group = ID)) +
#   geom_point(alpha=0.5,
#              size = 0.1) +
#   interior_annotation("b", sc_ltr_pos)

@@

```{r}
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
trimesh_removed_pbmc 
```
